---
title: "The Diamonds dataSet EDA"
author: "Ing. Ernesto Ibanez"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)

```

# The diamonds dataset EDA in R.

## It contains measurements on 10 different variables (like price, color, clarity, etc.) for 53,940 different diamonds [53940 10].

# This tutorial explains how to explore, summarize, and visualise the diamonds dataset in R.

# Load the diamonds data

```{r, echo=FALSE}
#load ggplot2 library
library(ggplot2)

diamonds
#view first six rows of diamonds dataset
head(diamonds)
names(diamonds)
#summarize diamonds dataset
summary(diamonds)
#display rows and columns
dim(diamonds)
str(diamonds)
```

# Visualise the diamonds Dataset

We can also create some plots to visualize the values in the dataset.

For example, we can create an histogram of the values for a certain variable:

```{r, echo=FALSE}
#create histogram of values for price
ggplot(data=diamonds, aes(x=price)) +
  geom_histogram(fill="steelblue", color="black",bins = 30) +
  ggtitle("Histogram of Diamods Price Values")

```

# Create a scatterplot of any pairwise combination of variables:

```{r, echo=FALSE}
#create scatterplot of carat vs. price, using cut as color variable
ggplot(data=diamonds, aes(x=carat, y=price, color=cut)) + 
  geom_point()

```

# Create scatterplot of price, grouped by cut

```{r, echo=FALSE}
ggplot(data=diamonds, aes(x=cut, y=price)) + 
  geom_boxplot(fill="steelblue")

```


```{r, echo=FALSE}
ggplot(diamonds, aes(color,price)) + geom_boxplot(fill="steelblue")
```


```{r, echo=FALSE}
ggplot(diamonds, aes(clarity,price)) + geom_boxplot(fill="steelblue")

```

# Plot histogram of diamond prices by cut

```{r, echo=FALSE}
by(diamonds$price, diamonds$cut, summary)
ggplot(data = diamonds, aes(x = price)) + 
  geom_histogram(binwidth = 250,fill="steelblue") + 
  facet_wrap(~cut, scales = "free_y") + 
  scale_x_continuous() + xlab('Price') + 
  ylab('Frequency')

```

#Plot histogram of diamond prices by carat

```{r, echo=FALSE}
diamonds$price_per_carat <- diamonds$price / diamonds$carat

ggplot(data = diamonds, aes(x = price_per_carat)) + 
  geom_histogram(binwidth = 250, fill="steelblue") + 
  facet_wrap(~cut, scales = "free_y") + 
  xlab('Price per Carat') + 
  ylab('Frequency')
```


# This session Summary

```{r, echo=FALSE}

sessionInfo()

```
