---
title: "The Iris Dataset EDA"
author: "Ing. Ernesto Ibanez"
date: "`r Sys.Date()`"
output:
  rmdformats::material:
    highlight: kate
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)

```

# The iris dataset is a dataset in R that contains measurements on 4 different attributes (in centimeters) for 50 flowers from 3 different species.

In 1935, Edgar Anderson collected data to quantify the geographic variations of iris flowers. The data set consists of 50 samples from each of the three sub-species ( iris setosa, iris virginica, and iris versicolor).

```{r}
# iris
head(iris)
str(iris)
attributes(iris)
summary(iris)
dim(iris)
# display column names
names(iris)
```

We can clearly see Iris dataSet is [150 5]. Species is the only one categorical variable

```{r}
df <- iris[, 1:4]
boxplot(df)
pairs(df)
stars(df)

PL <- df$Petal.Length
barplot(PL)
hist(PL)
SP <- iris$Species
pie(table(SP))
boxplot(PL ~ SP)
summary(aov(PL ~ SP))

PW <- df$Petal.Width
plot(PL, PW, col = SP)
abline(lm(PW ~ PL))


```
```{r}
summary(PL)

```
# This produces a lot of useful information about the distribution of petal length:

The minimum is 1.000, and the maximum is 6.900.

Average petal length is 3.758.

The mid-point, or median, is 4.350, as about half of the numbers are smaller than 4.350. Why the median is different from the mean? What happens if there is a typo and one number is entered 340cm instead of 3.40cm?

The 3rd quartile, or 75th percentile, is 5.100, as 75% of the flowers have petals shorter than 5.100. If a student’s GPA ranks 5th in a class of 25, he/she is at 80th percentile. The minimum is the 0th percentile and the maximum is the 100th percentile.

The 1st quartile, or 25th percentile, is 1.600. Only 25% of the flowers have petals shorter than 1.600. These summary statistics are graphically represented as a boxplot. Boxplots are more useful when multiple sets of numbers are compared.


```{r}
# View(iris) # show as a spreadsheet
SL <- df$Sepal.Length
hist(SL, main="Histogram of Sepal Length", xlab="Sepal Length",
     ylab='Frequency', xlim=c(4,8), col="blue", freq=FALSE)

SW <- df$Sepal.Width

hist(SW, main="Histogram of Sepal Width", xlab="Sepal Width",
     ylab='Frequency', xlim=c(2,5), col="darkorchid", freq=FALSE)
```


# Box Plot shows 5 statistically significant numbers- the minimum, the 25th percentile, the median, the 75th percentile and the maximum. It is thus useful for visualizing the spread of the data is and deriving inferences accordingly. 
```{r}
irisVer <- subset(iris, Species == "versicolor")
irisSet <- subset(iris, Species == "setosa")
irisVir <- subset(iris, Species == "virginica")
par(mfrow=c(1,3),mar=c(6,3,2,1))
boxplot(irisVer[,1:4], main="Versicolor, Rainbow Palette",ylim = c(0,8),las=2, col=rainbow(4))
boxplot(irisSet[,1:4], main="Setosa, Heat color Palette",ylim = c(0,8),las=2, col=heat.colors(4))
boxplot(irisVir[,1:4], main="Virginica, Topo colors Palette",ylim = c(0,8),las=2, col=topo.colors(4))
```

```{r}
# create boxplot of sepal width vs. sepal length
boxplot(Sepal.Length~Species,
        data=iris,
        main='Sepal Length by Species',
        xlab='Species',
        ylab='Sepal Length',
        col='steelblue',
        border='black')
        
# create scatterplot of sepal width vs. sepal length
plot(iris$Sepal.Width, iris$Sepal.Length,
     col='steelblue',
     main='Scatterplot',
     xlab='Sepal Width',
     ylab='Sepal Length',
     pch=19)

```

# Scatterplot Matrices
When we have two or more variables and we want to correlate between one variable and others so we use a scatterplot matrix.

pairs() function is used to create matrices of scatterplots.

```{r}
x <- iris[, 1:4]
pairs(x)
sessionInfo()
```
